\documentclass{article}
\title{Doxa: A Multi-Agent Simulation Platform for Economic and Social Research}
\author{Vincenzo et al.}
\date{\today}
\begin{document}
\maketitle





\begin{abstract}
Doxa is a modular, extensible platform for multi-agent simulation, integrating generative AI, economic modeling, and social dynamics. This paper presents the architecture, YAML-based scenario schema, agent logic, engine internals, API, and a detailed case study, providing a reproducible foundation for research in computational economics and AI-driven agent systems.
\end{abstract}

\section{Introduction}
Multi-agent simulation is a cornerstone of computational social science, economics, and artificial intelligence. Doxa addresses the need for a flexible, reproducible, and extensible platform that enables researchers to model complex agent interactions, market mechanisms, and world events, leveraging modern AI models for agent reasoning. This paper details the Doxa platform, its YAML-driven configuration, and its application to economic and social simulation.

\section{Related Work}
Agent-based modeling has a long tradition, with platforms such as NetLogo, MASON, and Repast. Recent advances integrate AI and machine learning for agent reasoning (e.g., OpenAI Gym, PettingZoo). Doxa distinguishes itself by combining declarative scenario modeling, generative AI integration, and a rich economic/market simulation toolkit, all accessible via a modern API.

\section{System Overview}
Doxa consists of a Python backend (DoxaEngine) and a web API server (FastAPI), exposing REST and WebSocket endpoints. Scenarios are specified in a single YAML file, which defines global rules, agent types, resources, markets, and world events. The engine parses this schema, instantiates agents, manages state, and orchestrates simulation steps. Agents are parameterized by persona, resource portfolio, behavioral rules, and can be linked to LLM providers (OpenAI, Google, Ollama, etc.).

\section{YAML Scenario Schema}
\label{sec:schema}
The scenario YAML file is the core of Doxa's reproducibility and flexibility. It defines:
\begin{itemize}
    \item \textbf{global\_rules}: Simulation-wide settings (timing, maintenance, kill/victory conditions, constraints, operations, markets, relations, dynamics).
    \item \textbf{actors}: List of agent types, each with identity, persona, initial portfolio, constraints, operations, trading mode, and economic parameters.
    \item \textbf{world\_events}: Optional list of scheduled or conditional events that affect resources, markets, or trust.
\end{itemize}
See Section~\ref{sec:case-study} for a concrete example (hormuz.yaml).

\section{Engine Architecture}
\label{sec:engine}

The Doxa engine is organized into modular subsystems, each responsible for a core aspect of the simulation. The main components are described below, with references to the codebase (\texttt{server/engine/}) and YAML schema fields.

\subsection{Economic Subsystem}
The economic logic of each agent is governed by the \texttt{AgentEconomics} module and the economic parameters defined in the YAML (e.g., utility function, risk aversion, discount factor). Agents can be risk-neutral or risk-averse (CRRA, CARA), and their decision-making incorporates liquidity preferences and price expectations. The economic subsystem ensures that agent actions (trades, operations) are consistent with their economic profile, and computes utility and risk metrics for analysis.

\subsection{Market Subsystem}
Markets are managed by the \texttt{MarketEngine} and \texttt{Market} classes. Doxa supports both over-the-counter (OTC) bilateral trades and centralized limit order books (LOBs) for each resource. Market configuration is fully declarative in YAML, including price bounds, clearing modes (per step, on order, call auction), and synthetic market makers (with configurable spread, depth, and inventory skew). The market subsystem matches orders, applies price impact, and enforces slippage and constraints, providing a realistic trading environment for agents.

\subsection{Relations Subsystem}
Social and trust relations are modeled as a directed, weighted graph managed by the \texttt{RelationGraph} and \texttt{RelationRecord} modules. Initial trust levels and relation types (ally, neutral, rival, enemy) are seeded from YAML. Trust dynamics are updated in response to agent actions (successful/rejected trades, broadcasts) and world events, with support for trust decay and contagion. Relations influence agent decision-making, negotiation, and the spread of events (e.g., panic contagion).

\subsection{Communication Subsystem}
Agents interact via private messages, public broadcasts, and negotiation protocols. The \texttt{DoxaAgent} class registers communication tools (send message, broadcast, propose trade, accept/reject trade) based on agent capabilities. Communication is event-driven and can trigger trust updates or world events. The communication subsystem is extensible, allowing custom tools to be defined in the YAML scenario.

\subsection{Resource Management}
Resources are the fundamental quantities tracked in agent portfolios and markets. The \texttt{SimulationEnvironment} maintains the state of all resources, applies maintenance costs, and enforces constraints (min/max bounds) at both global and agent levels. Resource transfers occur via operations, trades, market transactions, and world events. The engine ensures that all resource changes are validated and rolled back if constraints are violated.

\subsection{Events Subsystem}
World events are managed by the \texttt{WorldEventScheduler} and \texttt{WorldEventEffect} modules. Events can be scheduled (shocks, trends) or triggered by conditions (resource thresholds, agent states). Effects include resource changes, market price shifts, trust updates, and contagion. The event subsystem supports complex, multi-step event logic, enabling rich scenario evolution and exogenous shocks.

\subsection{API and Integration}
The API layer is implemented using FastAPI and exposes both REST and WebSocket endpoints. The REST API provides scenario management, simulation control, and data retrieval. The WebSocket endpoints (\texttt{/ws/agents}, \texttt{/ws/resources}) enable real-time streaming of agent actions, resource updates, and chat. The API is designed for integration with custom frontends, dashboards, and external controllers, supporting both synchronous and asynchronous workflows.

\section{Agent Model and Reasoning}
Agents in Doxa are autonomous, configurable entities. Each agent is defined by a persona, initial resources, constraints, and a set of operations. Agents can communicate, negotiate, trade (OTC and LOB), reason internally, and manage persistent vector memory (RAG via ChromaDB). Hierarchical organization (leaders and sub-agents) is supported. LLM integration enables advanced reasoning and language-based interaction.

\section{Market and Economic Simulation}
Doxa supports both OTC and limit order book (LOB) trading, with configurable market parameters (price bounds, clearing modes, market makers). Economic behavior is further shaped by agent-level utility functions, risk aversion, and liquidity preferences. The engine enforces constraints and rollbacks for infeasible operations.

\section{World Events and Dynamics}
World events can be scheduled (shocks, trends) or triggered by conditions (e.g., resource thresholds). Effects include resource changes, market price shifts, trust updates, and contagion. The YAML schema supports complex event logic, enabling rich scenario evolution.

\section{API and Integration}
The backend exposes a REST API and two WebSocket endpoints:\newline
\texttt{/ws/agents} for real-time agent actions, chat, and portfolio management;\newline
\texttt{/ws/resources} for real-time resource updates. This enables integration with custom frontends, dashboards, or external controllers.

\section{Scenario Design: Case Study}
\label{sec:case-study}
As a concrete example, Listing shows the YAML for the ``hormuz'' scenario, featuring two agent types (player, miners), two resources (gold, corn), and a set of world events and market rules.




\section{Experiments and Results}
% TODO: Describe experimental setup, metrics, and results. Include figures, tables, and analysis.

\section{Discussion and Limitations}
% TODO: Discuss strengths, extensibility, and current limitations. Suggest future work.

\section{Conclusion}
Doxa provides a robust, extensible platform for multi-agent simulation, integrating economic, social, and AI-driven reasoning. Its YAML-based configuration, modular engine, and API make it suitable for a wide range of research applications.

\section*{References}
% TODO: Add references to related work, foundational papers, and relevant documentation.

\end{document}
